---
title: "Comb-p on EWAS outputs"
author: "Fabien Kon-Sun-Tack, Florent Chuffart"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---

```{r, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
```

# Reading `ewas_mean`  outputs


```{r plot_ewas}
if (!exists("mreadRDS")) {
  mreadRDS = memoise::memoise(function(...) {print("reading..."); ret=readRDS(...); print("done."); return(ret)})
}
prefix = "ewas_mean"
ewas_mean = mreadRDS(paste0(prefix, ".rds"))
head(ewas_mean)

ewas = ewas_mean[,c("beta_C","beta_D","pval_DC.value")]
colnames(ewas) = c("beta_C", "beta_D", "lpv")
ewas$lpv = -log10(ewas$lpv)
layout(matrix(1:2, 1), respect=TRUE,)
plot(ewas[,"beta"], ewas[,"lpv"], main="Volcano plot", pch=".")
```

# Export EWAS for `Comb-p`

```{r build_bed}
range(ewas[,"lpv"])
range(ewas[,"beta"])
s = mreadRDS("../../data/study_hadaca_epic.rds")
pf = s$platform
head(ewas)
head(pf[,1:6])
lpval = ewas[,"lpv"]
names(lpval) = rownames(ewas)
# plot(lpval, main="Manhattan plot", pch=".")
```

Comb-p does not apreciate p-val lower that $10^{-45}$.
It produces `nan` values.
The following code cheats on this point.

```{r}
layout(matrix(1:2, 1), respect=TRUE)
plot(density(lpval), main="original pval")
# if (max(lpval)>40) {
#   lpval = 40*lpval/max(lpval)
# }
lpval[lpval > 30] = (15*(lpval[lpval>30]-30)/(max(lpval)-30))+30
# lpval[lpval > 40] = 40
plot(density(lpval), main="hacked pval")
```

```{r}
pval = 10^-lpval
# build and write bed file
bed = pf[rownames(ewas),c(1:3,5)] # Warning, here pf is global, it must be arg and indexed!!
head(bed)
bed$probes = names(pval)
bed$pval = pval
bed$beta_C = ewas[,"beta_C"]
bed$beta_D = ewas[,"beta_D"]
bed = bed[,c(1:3,5,6,7,8,4)]
colnames(bed) = c("chrom", "start", "end", "probes", "pval", "beta_C", "beta_D", "strand")

bed = na.omit(bed)
head(bed)
# write ewas res for combp
bed_ewas_filename = paste0("ewas4combp_", prefix,".bed")
bed[,1] = as.character(bed[,1])
bed = bed[order(bed[,1], bed[,2]),]
write.table(bed,file=bed_ewas_filename , sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
# bed = read.table(ewas_filename, sep="\t", header=TRUE)
```

# Run `Comb-p`

```{r combp}
pval_tresh = 10^-2
prefix2 = paste0(prefix, "_", pval_tresh)
# Run comb-p, run!!
cmd = "comb-p"
arg = paste0("pipeline -c 5 --seed ", pval_tresh, " --dist 1000 -p dmr_", prefix2," --region-filter-p 0.05 --region-filter-n 2 ", bed_ewas_filename)
print(paste(cmd, arg))
system2(cmd, arg)
```

# Results


Co-localization of Protein Z, Protein Z-Dependent protease inhibitor and coagulation factor X in human colon cancer tissue: implications for coagulation regulation on tumor cells.
Sierko E, Wojtukiewicz MZ, Zimnoch L, Tokajuk P, Ostrowska-Cichocka K, Kisiel W.
Thromb Res. 2012 Apr;129(4):e112-8. doi: 10.1016/j.thromres.2011.10.027. Epub 2011 Nov 17.
PMID: 22424030

 
Protein Z is present in human breast cancer tissue.
Sierko E, Wojtukiewicz MZ, Zimnoch L, Tokajuk P, Kisiel W.
Int J Hematol. 2011 May;93(5):681-683. doi: 10.1007/s12185-011-0846-3. Epub 2011 Apr 12.
PMID: 21479979 No abstract available.

Unbalance between plasma levels of Protein Z and protein Z-dependent inhibitor in patients with colorectal and pancreatic cancer: a pilot study.
Doat S, Borgel D, François JH, Bianchini E, Botton J, François D, Mitry E, Vasse M.
Thromb Res. 2014 Feb;133(2):299-300. doi: 10.1016/j.thromres.2013.11.015. Epub 2013 Nov 22.
PMID: 24315319 No abstract available.


Protein Z/protein Z-dependent protease inhibitor system in human non-small-cell lung cancer tissue.
Sierko E, Wojtukiewicz MZ, Zimnoch L, Ostrowska-Cichocka K, Tokajuk P, Ramlau R, Kisiel W.
Thromb Res. 2012 Apr;129(4):e92-6. doi: 10.1016/j.thromres.2011.09.005. Epub 2011 Oct 4.
PMID: 21975032

Protein Z/protein Z-dependent protease inhibitor system in loco in human gastric cancer.
Sierko E, Wojtukiewicz MZ, Zimnoch L, Tokajuk P, Ostrowska-Cichocka K, Kisiel W.
Ann Hematol. 2014 May;93(5):779-84. doi: 10.1007/s00277-013-1941-8. Epub 2013 Oct 26.
PMID: 24158387 Free PMC article.

Intron F G79A polymorphism of the protein Z gene in cancer patients with and without thrombosis.
Eroglu A, Ozturk A, Cam R, Akar N.
J Thromb Thrombolysis. 2009 Feb;27(2):204-6. doi: 10.1007/s11239-008-0199-9. Epub 2008 Feb 5.
PMID: 18246466


```{r results, eval=TRUE, results= "hold"}
# Read comb-p ewas input
bed_ewas = read.table(bed_ewas_filename, sep="\t", header=TRUE)
rownames(bed_ewas) = bed_ewas$probes
head(bed_ewas)
dim(bed_ewas)

# Read comb-p outputs
probes_file = paste0("dmr_", prefix2, ".fdr.bed.gz")
if (file.exists(probes_file)) {
  probes = read.table(gzfile(probes_file), comment="@", header=TRUE);
} else {
  warning(paste0("probes_file ", probes_file, " does not exist. Have to fix it! "))
  return(NULL)    
}
head(probes)
dim(probes)
if (sum( paste0(bed_ewas[,1], ":", bed_ewas[,2], "-", bed_ewas[,3]) !=  paste0(probes[,1], ":", probes[,2], "-", probes[,3])) != 0) {
  stop("Error when namimg probes.")
}
rownames(probes) = rownames(bed_ewas)

regions_file = paste0("dmr_", prefix2, ".regions-t.bed"   )
if (file.exists(regions_file)) {
  regions = read.table(regions_file, comment="@", header=TRUE, stringsAsFactors=FALSE)
} else {
  warning(paste0("regions_file ", regions_file, " does not exist. Have to fix it! "))
  return(NULL)
}

rownames(regions) = paste0(regions[,1], ":", regions[,2], "-", regions[,3])
head(regions[ rev(order(regions$n_probes)),],50)
bed_regions = regions
bed_regions[,4] = paste0(bed_regions[,1], ":", bed_regions[,2], "-", bed_regions[,3])
bed_regions[,5] = bed_regions[,"z_sidak_p"]
bed_regions[,6] = "+"
bed_regions = bed_regions[,1:6]
colnames(bed_regions) = c("chr", "start", "end", "name", "pval", "strand")
head(bed_regions)
new_regions = regions[ rev(order(regions$n_probes)),]


```

```{r eval=T, results="hold"}
# get probes
# idx_region = "chr12:114698194-114698504" #which(bed_regions[3]-bed_regions[2]==max(bed_regions[3]-bed_regions[2]))[1]
# idx_region = "chr6:32167275-32167620"
#
# chr17:48544347 48582736
# chr6 31568115 31582465
# gene = bed_regions[idx_region,]
gene = list("chr13", 113158210, 113158673, "PROZ", "1", "+")
# gene = list("chr11", 816000, 825500, "PNPLA2", "1", "+")
# gene = list("chr7", 27101178, 27145894, "HOXA", "1", "+")
# gene = list("chr6", 31559930, 31593095, "TNF", "1", "+")
# gene = list("chr4", 1192127, 1231942, "CTBP1", "1", "+")
# gene = list("chr14", 22400000, 22700000, "DAD1", "1", "+")
# gene = list("chr12", 124500726, 124528633, "NCOR2", "1", "+")

#gene = list("chr6", 33163968, 33164542, "COL11A2", "1", "+")
#gene = list("chr17", 864061, 864328 , "NXN", "1", "+")
#gene = list("chrX", 103214961, 103215296, "BEX4", "1", "+")
#gene = list("chr3", 50230303, 50230828, "GNAI2", "1", "+")
print(gene)
pf_meth = probes
pf_chr_colname = "X.chrom"
pf_pos_colname = "start"
up_str = 0
dwn_str = gene[[3]] - gene[[2]]
idx_probes = dmprocr::get_probe_names(gene, pf_meth, pf_chr_colname, pf_pos_colname, up_str, dwn_str)


layout(matrix(c(2, 2, 1, 1, 1, 1), 2), respect=TRUE)
ylim = c(0, min(max(-log10(probes[idx_probes[idx_probes%in%rownames(probes)],4:6])), 45))
plot(-log10(probes[idx_probes,]$p), col="grey", main=prefix2, ylim=ylim, pch=16)
points(-log10(probes[idx_probes,]$region.p), col=4, pch=16)
points(-log10(probes[idx_probes,]$region.q), col=2, pch=16)
legend("topright", legend=c("ewas pval", "slk pval", "slk qval"), col = c("grey", "blue", "red"), pch=16)

# plot it
layout(1)
plot(probes[idx_probes,]$start, -log10(probes[idx_probes,]$p), col="grey", main=prefix2, ylim=ylim, pch=16)
points(probes[idx_probes,]$start, -log10(probes[idx_probes,]$region.p), col=4, pch=16)
points(probes[idx_probes,]$start, -log10(probes[idx_probes,]$region.q), col=2, pch=16)
# segments(regions[idx_region,]$star, -log10(as.numeric(pval_tresh)), regions[idx_region,]$end, -log10(as.numeric(pval_tresh)), col=2)
legend("topright", legend=c("ewas pval", "slk pval", "slk qval"), col = c("grey", "blue", "red"), pch=16)




d = s$data
e = s$exp_grp

idx_sample = rownames(e)[substr(rownames(e),1, 1)%in% c("C", "D")]

par(mar=c(10, 4.1, 4.1, 2.1))
layout(matrix(c(1, 1, 2, 2, 2, 2), 2), respect=TRUE)
plot(as.numeric(as.factor(substr(idx_sample,1, 1))), 1:length(idx_sample)) 
colors = c("cyan", "black", "red")
cols = colorRampPalette(colors)(20)
breaks = seq(0, 1, length.out = length(cols) + 1)
main = gene[[4]]
image(d[idx_probes,idx_sample], col=cols, breaks=breaks, xaxt="n", yaxt="n", main=main)
axis(1, (1:nrow(d[idx_probes,idx_sample]) - 1)/(nrow(d[idx_probes,idx_sample]) - 1), rownames(d[idx_probes,idx_sample]), las = 2)

```

```{r echo=T, eval=T}

DMR = apply(new_regions,1,function(l){
  chr = l[[1]]
  st = as.numeric(l[[2]])
  stop = as.numeric(l[[3]])
  r1 = subset(bed_ewas, chrom == chr & start >= st & end <= stop)
  ret = c(
      min_beta_C = min(r1$beta_C),
      max_beta_C = max(r1$beta_C),
      mean_beta_C = mean(r1$beta_C),
      min_beta_D = min(r1$beta_D),
      max_beta_D = max(r1$beta_D),
      mean_beta_D = mean(r1$beta_D)
  )
  return(ret)
})
DMR = t(DMR)
new_regions = cbind(new_regions,DMR)


```

```{r ggplot, echo=F, eval=T}
library(reshape2)
library(ggplot2)
df = melt(d[idx_probes,idx_sample], id.vars = colnames(d[idx_probes,idx_sample]))
names(df) = c("probes", "conditions", "value")
df$cond = substr(df$conditions,1,1)
plot = ggplot(df, aes(x=probes, y=value, color=cond)) + geom_point()
plot

boxplot(value~cond+probes,df, las=2,border=c(2,4), ylim=c(0,1))

```
```{r test, eval=T, results="verbatim"}
idx_C = colnames(d)[17:24]
idx_D = colnames(d)[25:32]
don = na.omit(d)
test = apply_func(don[idx_probes,], 1, function(l){
  #print(l)
  #l = don[idx_probes[1],]
  data = e[colnames(don),]
  data$meth = l
  m_DC = lm(meth~cond, data[c(idx_D, idx_C),])
  m_DC$coef
  data[idx_C,]
  boxplot(meth~cond, data[c(idx_D, idx_C),])
  abline(h=mean(data[idx_C,3]), col=4)
  abline(h=mean(data[idx_D,3]), col=2)
  abline(h=m_DC$coef[[1]], lty=2)
  abline(h=m_DC$coef[[1]]+m_DC$coef[[2]], lty=3)

}
)

```
# Session Information

```{r results="verbatim"}
sessionInfo()
```

